Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-05-26 cvx-flats-neurips21/cvx-flats.tex
section 4

\raf{New intro... something along these lines}

We now turn to two main applications of Theorem~\ref{thm:convex-flats-inf-dim}: new lower bounds on the convex consistency dimension of risk measures (\S~\ref{sec:risk-measures}), and the mode and modal interval (\S~\ref{sec:mode-modal-interval}).
In both cases, we build on previous results due to \citet{frongillo2020elicitation} and \citet{dearborn2020indirect} which showed lower bounds with respect to \emph{identifiable} properties;
a property is $d$-identifiable if its level sets are all $d$-flats, as in Figure~\ref{fig:quadrants-simplices} (Q3).
In contrast, properties elicited by non-smooth convex losses are generally not identifiable.
For example, the properties elicited by hinge loss and the abstain surrogate are not identifiable, as their level sets are not flats; see Figure~\ref{fig:quadrants-simplices} (Q1).
It therefore might appear that entirely new ideas are needed.
Indeed, both papers above pose developing similar bounds with respect to convex-elicitable properties as a major open question.

Using our framework, we establish such bounds, resolving both open questions.
The central observation, from Theorem~\ref{thm:convex-flats-inf-dim}, is that the level sets of $d$-dimensional convex-elicitable properties, if not $d$-flats themselves, are \emph{unions} of $d$-flats.
Using this observation, we can modify the arguments of~\citet{frongillo2020elicitation} and~\citet{dearborn2020indirect} to give similar bounds.
In all cases, our lower bounds match the identifiable case, though we conjecture they are not tight in all cases.

\raf{MOVE}

\citet{frongillo2020elicitation} rely on finding a dimension $d$ such that the level sets of certain risk measures $\gamma$ have too much curvature to contain any $d$-flat.
Thus, the elicitation complexity with respect to identifiable properties is greater than $d$.
In particular, we recover their main lower bound for the large class of Bayes risks.

\raf{OLD intro}
%commented out Jun 1 2020
%Theorem~\ref{thm:consistent-implies-indir-elic} allows us to use convex elicitation complexity as a tool to understand efficiency of consistent convex surrogates for a given property, which is more often what is given in a continuous estimation setting.
%For example, when one wants to learn an $\alpha$-quantile, we start with the property rather than a loss.
%In the literature, pinball loss $L(r,y) = (r-y)(\ones_{r \geq y} - \alpha)$ typically appears without explanation or justification as to why it is relevant in relation to the $\alpha$-quantile.
%Elicitation teaches us why pinball loss is consistent for learning a quantile as the pinball loss elicits the $\alpha$-quantile.
In continuous estimation problems, often one is not given a target loss, but instead a target (conditional) statistic of the data one wishes to estimate.
Examples include estimating the mean or variance of $y$ conditioned on a given $x$.
In this setting, Theorem~\ref{thm:convex-flats-inf-dim} gives lower bounds on the prediction dimension of convex losses with a link to the desired conditional statistic, i.e., the convex elicitation complexity.
In particular,
% following an argument from~\citet{frongillo2020elicitation},
Theorem~\ref{thm:bayes-risk-lower-bound} below yields new bounds on the convex elicitation complexity of statistics which quantify risk or uncertainty such as variance, entropy, or financial risk measures.

These bounds address an open question of~\citet{frongillo2020elicitation}, that of developing a theory of elicitation complexity with respect to convex-elicitable properties.
The lower bounds of previous work are essentially all with respect to \emph{identifiable} properties;
a property is $d$-identifiable if its level sets are all $d$-flats.
\citet{frongillo2020elicitation} rely on finding a dimension $d$ such that the level sets of certain risk measures $\gamma$ have too much curvature to contain any $d$-flat.
Thus, the elicitation complexity with respect to identifiable properties is greater than $d$.

In contrast, properties elicited by non-smooth convex losses are generally not identifiable.
For example, the properties elicited by hinge loss and the abstain surrogate are not identifiable, as their level sets are not flats (see Figure~\ref{fig:fsd-flats-abstain}).
It therefore might appear that entirely new ideas are needed.
Our framework is closely related to identifiability, however; Theorem~\ref{thm:convex-flats-inf-dim} states that the level sets of $d$-dimensional convex-elicitable properties, if not $d$-flats themselves, are unions of $d$-flats.
Thus, the general logic of~\citet{frongillo2020elicitation} can still apply.
In particular, we recover their main lower bound for the large class of Bayes risks.
\btw{Old version for finite $\Y$ is in the NeurIPS 2020 submission; especially lots more detail on variance / norm / entropy examples}

\subsection{Risk measures}
\label{sec:risk-measures}

@@

We now illustrate the theorem's application to conditional value at risk.
Several other applications from \citet{frongillo2020elicitation}, such as risk measures, entropy, and norms, follow similarly.

\paragraph{Conditional Value at Risk.}

\citet{frongillo2020elicitation} observe that one of the most prominent financial risk measures, the conditional value at risk (CVaR), can be expressed as a Bayes risk.
In particular, for $0 < \alpha < 1$, we may define
\begin{align}
  \CVaR_\alpha(p)
  &= \inf_{r\in\reals} \E_p\left\{ \tfrac 1 \alpha(r-Y)\ones_{r\geq Y}- r\right\}~,
  \label{eq:elic-complex-es-3}
\end{align}
which is the Bayes risk of the transformed pinball loss $L_\alpha(r,y) = \tfrac 1 \alpha(r-y)\ones_{r\geq y}-r$.
In turn, $L_\alpha$ elicits the $\alpha$-quantile, the quantity $q_\alpha(p)$ such that $\Pr_p[Y \geq q_\alpha(p)] = \alpha$.
Following \citet{frongillo2020elicitation}, we will restrict to the set $\P_q$ of probability measures over $\reals$ with connected support and whose CDFs are strictly increasing on their support, so that $q_\alpha$ is single-valued.
Under mild assumptions, we find that there is no consistent real-valued convex surrogate for $\CVaR_\alpha$.

>>>>

We now turn to two main applications of Theorem~\ref{thm:convex-flats-inf-dim}: new lower bounds on the convex consistency dimension of risk measures (\S~\ref{sec:risk-measures}), and the mode and modal interval (\S~\ref{sec:mode-modal-interval}).
In both cases, we build on previous results due to \citet{frongillo2020elicitation} and \citet{dearborn2020indirect} which showed lower bounds with respect to \emph{identifiable} properties;
a property is $d$-identifiable if its level sets are all $d$-flats, as in Figure~\ref{fig:quadrants-simplices} (Q3).
In contrast, properties elicited by convex losses are generally not identifiable, particularly when the loss is non-smooth.
For example, the properties elicited by hinge loss and the abstain surrogate are not identifiable, as their level sets are not flats; see Figure~\ref{fig:quadrants-simplices} (Q1).
It therefore might appear that entirely new ideas are needed.
Indeed, both papers above pose developing similar bounds with respect to convex-elicitable properties as a major open question.

Using our $d$-flats framework, we establish such bounds, thus resolving both open questions.
The central observation is that the level sets of $d$-dimensional convex-elicitable properties, if not $d$-flats themselves, are \emph{unions} of $d$-flats from Theorem~\ref{thm:convex-flats-inf-dim}.
It turns out that the arguments of~\citet{frongillo2020elicitation} and~\citet{dearborn2020indirect} essentially rely only on the containment of $d$-flats in other level sets, rather than the full structure of identifiable properties.
As such, although quite subtle in the case of risk measures, we can modify their arguments for convex-elicitable properties: since no $d$-flat could be contained in a particular level set, no union of $d$-flats could be either.
Our lower bounds therefore match both of these papers, though we conjecture that the convex consistency dimension is strictly higher in some cases.

\btw{OLD intro commented out}
% In continuous estimation problems, often one is not given a target loss, but instead a target (conditional) statistic of the data one wishes to estimate.
% Examples include estimating the mean or variance of $y$ conditioned on a given $x$.
% In this setting, Theorem~\ref{thm:convex-flats-inf-dim} gives lower bounds on the prediction dimension of convex losses with a link to the desired conditional statistic, i.e., the convex elicitation complexity.
% In particular,
% % following an argument from~\citet{frongillo2020elicitation},
% Theorem~\ref{thm:bayes-risk-lower-bound} below yields new bounds on the convex elicitation complexity of statistics which quantify risk or uncertainty such as variance, entropy, or financial risk measures.

% These bounds address an open question of~\citet{frongillo2020elicitation}, that of developing a theory of elicitation complexity with respect to convex-elicitable properties.
% The lower bounds of previous work are essentially all with respect to \emph{identifiable} properties;
% a property is $d$-identifiable if its level sets are all $d$-flats.
% \citet{frongillo2020elicitation} rely on finding a dimension $d$ such that the level sets of certain risk measures $\gamma$ have too much curvature to contain any $d$-flat.
% Thus, the elicitation complexity with respect to identifiable properties is greater than $d$.

% In contrast, properties elicited by non-smooth convex losses are generally not identifiable.
% For example, the properties elicited by hinge loss and the abstain surrogate are not identifiable, as their level sets are not flats (see Figure~\ref{fig:fsd-flats-abstain}).
% It therefore might appear that entirely new ideas are needed.
% Our framework is closely related to identifiability, however; Theorem~\ref{thm:convex-flats-inf-dim} states that the level sets of $d$-dimensional convex-elicitable properties, if not $d$-flats themselves, are unions of $d$-flats.
% Thus, the general logic of~\citet{frongillo2020elicitation} can still apply.
% In particular, we recover their main lower bound for the large class of Bayes risks.
% \btw{Old version for finite $\Y$ is in the NeurIPS 2020 submission; especially lots more detail on variance / norm / entropy examples}

\subsection{Risk measures (Q4)}
\label{sec:risk-measures}

The problem of estimating a risk or uncertainty measure of $Y$ has applications in financial regulation~\cite{acerbi2014back-testing,fissler2016expected,campbell2005review}, robust engineering design\raft{add cites from AFOSR or EnvAI}, and algorithmic fairness\raft{Bob}.
For examples of risk measures, one may wish to estimate the upper confidence bound $\E[Y] + \lambda \Var[Y]$, or the Conditional Value at Risk (CVaR) defined below, in either conditional or unconditional contexts.
Uncertainty measures can be unbiased versions of the above, such as the variance, or an entropy or norm of the distribution of $Y$.
Risk and uncertainty measures are typically not elicitable, so this problem falls under Quadrant 4.
\citet{frongillo2020elicitation} give prediction dimension lower bounds for a broad class of risk and uncertainty measures, namely Bayes risks.
As stated above, these bounds are with respect to identifiable properties, and bounds for convex surrogates are left as a major open question.

We resolve this open question using our $d$-flats framework, giving a matching result (Theorem~\ref{thm:bayes-risk-lower-bound}) for convex-elicitable properties.
% \citet{frongillo2020elicitation} rely on finding a dimension $d$ such that the level sets of certain risk measures $\gamma$ have too much curvature to contain any $d$-flat.
% With a careful modification of their argument, we can extend this observation to unions of $d$-flats.
First we recall the definition of the Bayes risk.

@@

To illustrate the theorem, we briefly apply it to one of the most prominent financial risk measures, the conditional value at risk (CVaR).
Several other applications from \citet{frongillo2020elicitation}, such as other risk measures, entropy, and norms, follow similarly.
The authors observe that CVaR can be expressed as a Bayes risk; for $0 < \alpha < 1$, we may define
\begin{align}
  \CVaR_\alpha(p)
  &= \inf_{r\in\reals} \E_p\left\{ \tfrac 1 \alpha(r-Y)\ones_{r\geq Y}- r\right\}~,
  \label{eq:elic-complex-es-3}
\end{align}
which is the Bayes risk of the transformed pinball loss $L_\alpha(r,y) = \tfrac 1 \alpha(r-y)\ones_{r\geq y}-r$.
In turn, $L_\alpha$ elicits the $\alpha$-quantile, the quantity $q_\alpha(p)$ such that $\Pr_p[Y \geq q_\alpha(p)] = \alpha$.
Following \citet{frongillo2020elicitation}, we will restrict to the set $\P_q$ of probability measures over $\reals$ with connected support and whose CDFs are strictly increasing on their support, so that $q_\alpha$ is single-valued.
Under mild assumptions, we find that there is no consistent real-valued convex surrogate for $\CVaR_\alpha$.
